Thermodynamic fragments for magic states in quantum computation

### 2021-03-31 magic_paper.tex
minor additions
Cleaned up after Sec.II

\begin{document}
% Distillation rate vs bound
% Proofs in App. A
% Eq. numbering in last line always
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds
% Sec: Intro
% Sec: Abstract
% Figs consistency (esp. figs 3,4,5,6 and 1,7) + Figure 0 ?
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Central to our construction is the representation of any quantum state or quantum operation on a system of dimension $d$ in terms of quasi-probability representations on a discrete phase space~\cite{Ferrie_2008}. This construction is a discrete version of Wigner representations in quantum optics.

@@

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmx}: k \in \mathbb{Z}_d, \bmx \in \cal{P}_d^n\}.
\end{equation}
The Clifford operations $ \cal{C}_d^n $ are then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabiliser states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3}. Finally, we define $\stab$ as the convex hull of all pure stabiliser states, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is some Clifford unitary.

@@

The phase-point operators form an orthogonal (with respect to the Hilbert-Schmidt inner product) operator basis as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them,
\begin{equation}
    \rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz},
\end{equation}
where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmx)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$ (see~\cref{app:wigner} for details). More precisely, $W_\rho(\x)$ is a real-valued function on $\P_d$ with the property that $\sum_{\x} W_\rho(\bmx) = 1$.  In~\cref{fig:wstate_examples}, we show Wigner distributions of different types of qutrit states.

\begin{figure}%
    \centering
    \subfigure[][]{%
    \label{fig:maxmix}%
    \includegraphics[height=2cm]{figs/maxmixed.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:zero}%
    \includegraphics[height=2cm]{figs/zerostate.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:bound}%
    \includegraphics[height=2cm]{figs/boundstate.pdf}
    %\caption{Bound state}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:strange}%
    \includegraphics[height=2cm]{figs/strangestate.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Qutrit Wigner distributions of varying magic.} 
    \subref{fig:maxmix} Maximally mixed state $\frac{1}{3}\id$; \subref{fig:zero} Stabilizer zero state $\ketbra{0}{0}$; \subref{fig:bound} A non-stabiliser Wigner-positive state; \subref{fig:strange} Magic Strange state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$, coined in~\cite{cit:veitch2}.
    }%
    \label{fig:wstate_examples}
\end{figure}

@@

The state-injection model for quantum computation~\cite{cit:bravyi} is a key avenue to realising a full-scale, fault-tolerant quantum computer. Within this approach, Clifford unitaries can be done in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, it is impossible to have a universal set of transversal gates and so while Clifford unitaries can be realised transversally one must find ways around the Eastin-Knill restriction. The state-injection approach supplements the Clifford gates with noisy quantum states, called `magic states', that are outside the set of stabliser states and, when incorporated into Clifford circuits, provide universality. The obstacle to this is that the states are invariably noisy and so protocols must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates. A central question is then: given a number $n$ of noisy magic states how many purified magic states can we obtain within the fault-tolerant set of quantum operations?

To address such questions both concrete distillation protocols have been developed, and there has been analysis of bounds of distillation rates within natural theoretical frameworks. These frameworks provide theories of magic in which one views magic states as ``resource'' states with respect to a natural class of quantum operations that are consider cheap, or ``free''~\cite{Gour_2019}. One very natural class of free operations considered are those obtained from Clifford operations, measurements and the ability to discard quantum systems. However there are several other candidate frameworks.

For any theory of magic a natural route to bounding the distillation rates obtainable is through the concept of a \emph{magic monotone}. A magic monotone is a real-valued function of any quantum state $\M(\rho)$ that is monotonically non-increasing under the free operations of the magic theory. More precisely $\M(\sigma) \le \M(\rho)$ whenever it is possible to convert $\rho$ into $\sigma$ using at least one of the free operations available (e.g. Clifford operations).

@@

A range of monotones have been developed in the literature~\cite{cit:howard, Wang_2020, Seddon_2021} and each provides a means to upper bound distillation rates within a particular theory of magic. In this work we shall consider a broad framework for magic, with conditions that any theory of magic should satisfy. However, in contrast to prior works we shall not construct magic monotones, but instead develop a generalisation to resource monotones. The approach we take will also allow us to address how distillation rates are constrained by physical properties -- such as a system having biased-noise, some equilibrium structure or operations that manifest invariances, such as axial symmetry around the qubit $Z$-axis.

@@

Within the Wigner representation all negatively represented states are magic states, although in general the converse is not true. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, and therefore it is natural to consider focus on describing those negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, and so the particular choice of free states is not so important for our analysis. The remaining component that needs addressing is the set of quantum operations that is considered free, or easy. Stabilizer operations are the most natural set of operations as they admit a classically efficient description for uniform quantum circuits. However, other choices of operations have been considered in the literature~\cite{cit:ahmadi, cit:seddon, Wang_2019}. Common to these theories is that free $\E$  is such that $(\id \otimes \E) (\rho)$ has a positive Wigner distribution whenever $ \rho$ has a positive Wigner distribution. \nick{This is not true for stabilizer-preserving operations~\cite{cit:ahmadi}, since ``completely'' stabilizer-preserving operations~\cite{cit:seddon} is a strict subset} Therefore, the condition that $\E$ does not generate Wigner negativity is an essential requirement of any magic theory.

Any quantum channel admits a Wigner representation, in terms of the phase space operators defined earlier. If $\E$ is some quantum channel from a quantum system $A$ to a quantum system $B$, and $\J(\E)$ is its associated Choi state then we can define
\begin{equation}
W_{\E}(\y |\x) := d_A^2 W_{\J(\E)}(\bar{\x} \oplus \y),
\end{equation}
where $\bar{x} =(x_1, -p_1, \dots , x_n, -p_n)$ can be viewed as the time-reversed version of $\x$ in the discrete phase space, where momenta are reflected while coordinates remain unchanged.

With this representation of a quantum channel, we can view $W_{\E}( \y | \x)$ as a transition matrix sending $\x \rightarrow \y$, which is consistent with the representation of quantum states, as
\begin{equation}
W_{\E(\rho)} (\y) = \sum_{\x \in \P_d} W_{\E}( \y | \x) W_\rho(\x),
\end{equation}
for any $\E$ and $\rho$. Now if $\E$ sends any positively represented quantum state to another positively represented quantum state then it can be shown~\cite{Wang_2019} that $W_{\E}(\y |\x)$ must form a stochastic matrix. In particular, all Clifford operations correspond to stochastic matrices in this Wigner representation. However, it is important to note that not all stochastic maps on the phase space correspond to valid quantum operations. In particular, the stochastic maps must respect the symplectic structure of the phase space and so this provides an additional non-trivial constraint.

In what follows we shall therefore assume that we have a magic theory $\R = (\F, \O)$ in which the free states $\F$ form a closed, convex set with each state represented by a positive Wigner function, while the free operations $\O$ are represented by stochastic maps acting on the discrete phase space. \nick{Should we justify closedness, i.e. we require state positivity and Wigner-positivity both of which are closed linear constraints. I think we don't require convexity, we just need boundedness which trivially holds for quantum states}\ddd{[Fine print. No issue.]}

@@

\begin{figure}
    \centering
    \includegraphics[height=4.5cm]{figs/lctoy.pdf}
    \caption{\textbf{Quasi-probability Lorenz curve comparison}.
    The Lorenz curves are constructed by the Wigner distributions illustrated in~\cref{fig:wstate_examples} in the unital fragment $\O_{\id/d}$ for $d=3$.
    The maximally mixed state curve is simply the line connecting $(0,0)$ and $(1,1)$.
    There is no operation in $\O_{\id/d}$ that can convert $\ket{\rm{S}}$ to $\ket{0}$, as their Lorenz curves intersect.\ddd{[Is debatable if this figure gives much...is a bit boring.]}
    }
    \label{fig:lctoy}
\end{figure}

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> := \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. The reason for this particular choice is that the Wigner distribution is very simple. Specifically $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the value $1/6$ at all other points. We also define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq \left[ (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id \right],
\end{equation}
where $\epsilon$ is the depolarization noise parameter. in $\O_\sigma$, where the pure magic state $\ket{\rm{S}}$ is induced with depolarising noise. In fact, any magic state $\rho$ can be processed via Clifford operations and put into this canonical form for some $\epsilon \ge 0$.

@@

Note that the last elbow bounds have a dependence on $n$, which interestingly is due to $n$ being even.
Had we considered \textbf{odd} number of copies $n$, then the first elbow bound would depend on $n$ and the last elbow bound would not.

>>>>

\begin{document}
% Distillation rate vs bound
% Proofs in App. A
% Eq. numbering in last line always
% Sec: App. D
% Sec: General resources
% Sec: Lower bounds
% Sec: Intro
% Sec: Abstract
% Figs consistency (esp. figs 3,4,5 and 1,6) + Figure 0 ?
% Spacing (fig positioning, sections, paragraphs, QED symbols)

@@

Central to our construction is the representation of any quantum state or quantum operation on a system of dimension $d$ in terms of quasi-probability representations on a discrete phase space~\cite{Ferrie_2008}. This construction is a discrete version of Wigner representations in quantum optics~\cite{Wigner_1932, Vourdas_2004}.

@@

The displacement operators form the Heisenberg-Weyl group~\cite{Folland_1989, Bengtsson_2006} under matrix multiplication modulo phases,
\begin{equation}\label{eq:gp}
    {\rm{HW}}_d^n \coloneqq \{ \tau^k D_{\bmx}: k \in \mathbb{Z}_d, \bmx \in \cal{P}_d^n\}.
\end{equation}
The Clifford operations $ \cal{C}_d^n $ are then defined as the set of unitaries that normalise the Heisenberg-Weyl group~\cite{Appleby_2005}. We may define the pure stabiliser states as those states obtained by acting on $|0\>$ with Clifford unitaries~\cite{cit:gross3} and, finally, the full set of stabilizer states as the convex hull of all pure stabilisers, namely all probabilistic mixtures of states of the form $U|0\>\<0|U^\dagger$ where $U$ is Clifford.

@@

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them,
\begin{equation}
    \rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz},
\end{equation}
where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmx)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\x)$ is a bounded real-valued function on $\P_d$ with the normalisation property $\sum_{\x} W_\rho(\bmx) = 1$ (see~\cref{app:wigner} for details).  In~\cref{fig:wstate_examples}, we show Wigner distributions of different types of qutrit states.

\begin{figure}[t]
    \centering
    \subfigure[][]{%
    \label{fig:maxmix}%
    \includegraphics[height=2cm]{figs/maxmixed.pdf}
    %\caption{Maximally mixed state $\frac{1}{3}\id$}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:zero}%
    \includegraphics[height=2cm]{figs/zerostate.pdf}
    %\caption{Zero state $\ketbra{0}{0}$}%
    }\\
    \subfigure[][]{%
    \label{fig:bound}%
    \includegraphics[height=2cm]{figs/boundstate.pdf}
    %\caption{Bound state}%
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:strange}%
    \includegraphics[height=2cm]{figs/strangestate.pdf}
    %\caption{Strange state $\ketbra{S}{S}$}%
    }
    \caption{\textbf{Qutrit Wigner distributions of varying magic.} 
    \subref{fig:maxmix} Maximally mixed state $\frac{1}{3}\id$; \subref{fig:zero} Stabilizer zero state $\ketbra{0}{0}$; \subref{fig:bound} A non-stabiliser Wigner-positive state; \subref{fig:strange} Magic Strange state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$, coined in~\cite{cit:veitch2}.
    }%
    \label{fig:wstate_examples}
\end{figure}

Any quantum operation $\E$ can also admit a Wigner representation. If $\E$ maps some quantum system $A$ to a quantum system $B$, and $\J(\E)$ is its associated Choi state then we can define
\begin{equation}
W_{\E}(\y |\x) \coloneqq d_A^2 W_{\J(\E)}(\bar{\x} \oplus \y),
\end{equation}
where $\bar{x} =(x_1, -p_1, \dots , x_n, -p_n)$ can be viewed as the time-reversed version of $\x$ in the discrete phase space, where momenta are reflected while position coordinates remain unchanged.

@@

The state-injection model for quantum computation~\cite{cit:bravyi} is a key avenue to realising a full-scale, fault-tolerant quantum computer. Within this approach, Clifford unitaries can be implemented in a robust, fault-tolerant way. However, due to the Eastin-Knill theorem~\cite{Eastin_2009}, it is impossible to have a universal set of transversal gates, and although Clifford unitaries can be realised transversally, one needs to find ways around the Eastin-Knill restriction. The state-injection approach supplements the Clifford unitaries with noisy quantum states, called `magic states', that are outside the set of stabliser states and, when incorporated into Clifford circuits, provide universality. The obstacle to this is that the states are invariably noisy and so protocols must be employed to purify many copies of the magic states and improve the overall performance of the induced quantum gates. A central question is then: given a number $n$ of noisy magic states how many purified magic states can we obtain within the fault-tolerant set of quantum operations?

To address such questions, concrete distillation protocols have been developed, and there has also been analysis of bounds on distillation rates within natural theoretical frameworks. These frameworks provide theories of magic in which one views magic states as ``resource'' states with respect to a natural class of quantum operations that are consider cheap, or ``free''~\cite{Gour_2019}. One very natural class of free operations considered are those obtained from Clifford operations, measurements and the ability to discard quantum systems. However, there are several other candidate frameworks~\cite{cit:ahmadi, cit:seddon, Wang_2019}.

In any theory of magic, a natural route to bounding the distillation rates obtainable is through the concept of a \emph{magic monotone}. A magic monotone is a real-valued function of any quantum state $\M(\rho)$ that is monotonically non-increasing under the free operations of the magic theory. More precisely $\M(\sigma) \le \M(\rho)$ whenever it is possible to convert $\rho$ into $\sigma$ using free operations (e.g. Clifford operations).

@@

A range of monotones have been developed in the literature~\cite{cit:howard, Wang_2020, Seddon_2021} and each provides a means to upper bound distillation rates within a particular theory of magic. In this work we shall consider a broad framework for magic, with conditions that any magic theory should satisfy. However, in contrast to prior works we shall not construct magic monotones, but instead develop a generalisation to resource monotones. The approach we take will also allow us to address how distillation rates are constrained by physical properties -- such as a system having biased noise, some equilibrium structure or operations that manifest invariances, such as axial symmetry around the qubit $Z$-axis.

@@

Within the Wigner representation, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:campbell}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed (free) quantum operations, i.e. the ones that are considered free. The assumption we require on free operations is simply that they send any free state to another free state,meaning that $\F$ must be closed, which is a common assumption among all magic theories.

\nick{``the condition that $\E$ does not generate Wigner negativity is an essential requirement of any magic theory'': this is not true for stabilizer-preserving operations (SPOs)~\cite{cit:ahmadi}. Note that ``completely'' SPOs~\cite{cit:seddon} are a strict subset. However, SPOs are a subset of Wang's CPWP operations~\cite{Wang_2019} so our results hold for SPOs too.}

We can view the Wigner representation $W_{\E}( \y | \x)$ of a quantum channel $\E$ as a transition matrix mapping points $\x \rightarrow \y$, which is consistent with the representation of quantum states, as
\begin{equation}
	W_{\E(\rho)} (\y) = \sum_{\x \in \P_d} W_{\E}( \y | \x) W_\rho(\x),
\end{equation}
for any $\E$ and $\rho$. Now if $\E$ sends all positively represented quantum state to other positively represented quantum states then it can be shown~\cite{Wang_2019} that $W_{\E}(\y |\x)$ must form a stochastic matrix. In particular, all Clifford operations correspond to stochastic matrices in this Wigner representation. However, it is important to note that not all stochastic maps on the phase space correspond to valid quantum operations. In particular, the stochastic maps must respect the symplectic structure of the phase space, which is an additional non-trivial constraint.

In what follows we shall therefore assume that we have a magic theory $\R = (\F, \O)$ in which the free states $\F$ form a closed set with each state represented by a non-negative Wigner function, while the free operations $\O$ are represented by stochastic maps acting on the discrete phase space.

@@

@@

For our distillation protocols we need to choose a canonical pure magic state as a target. For this we shall choose the Strange state $|S\>$ given as
\begin{equation}
|S\> := \frac{1}{\sqrt{2}} (|1\> + |2\>),
\end{equation}
in the computational basis. The reason for this particular choice is that the Wigner distribution is very simple. Specifically $W_S(\x)$ has a single negative value of $-1/3$ at $\x =(0,0)$ and the value $1/6$ at all other points. We also define the $\epsilon$--noisy Strange state as
\begin{equation}\label{eq:noisysn}
    \rho_{\rm{S}}(\epsilon)\coloneqq (1 - \epsilon) \ket{\rm{S}}\bra{\rm{S}} + \epsilon \frac{1}{3}\id,
\end{equation}
where $\epsilon$ is the depolarization noise parameter. in $\O_\sigma$, where the pure magic state $\ket{\rm{S}}$ is induced with depolarising noise. In fact, any magic state $\rho$ can be processed via Clifford operations and put into this canonical form for some $\epsilon \ge 0$.

@@

Note that the last elbow bounds have a dependence on $n$, which interestingly is due to $n$ being even.
Had we considered \textbf{odd} number of copies $n$, then the first elbow bound would depend on $n$ and the last elbow bound would not.
This, along with the unital fragment analysis, makes me think that calculating the bounds again for odd $n$ or $\epsilon > 3/7$ would reveal nice symmetries of the analysis, but it's probably not worth the time, so including last elbow analysis properly feels incomplete.
